---
title: "Analysis_of_experimental_results"
author: "Ludvig Olsen"
date: "10/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Attach libraries
```{r}
library(tidyverse)
library(knitr)

data_path <- "results_data/"

```

How to save high resolution images:
```{r}

standard_width = 5
standard_height = 4
standard_res = 300

# Testing tiff_saver
# tiff("x.tiff", units="in", width=5, height=4, res=300)
# data.frame("x"=c(1,2,3,4,5), "y"=c(2,4,6,8,10)) %>% 
#   ggplot(aes(x,y)) +
#   geom_line()
# dev.off()
  

```


## Load data

### Condition: Increasing both

In this condition the number of migrants go up over time, while the epoch size becomes smaller (so migrations/communications become more frequent).

```{r}
data_timeseries_increasing_both <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsIncreasingMigrationBoth.txt")) %>% 
  mutate(condition = "Increasing_both",
         migration_direction = "Increasing") 

data_results_increasing_both <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsFinalResultsIncreasingMigrationBoth.txt")) %>% 
  mutate(condition = "Increasing_both",
         migration_direction = "Increasing")

```

### Condition: Increasing number of migrants

Increase the number of individuals that migrate after each epoch.

```{r}
data_timeseries_increasing_migrants <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsIncreasingMigrationMigrants.txt")) %>% 
  mutate(condition = "Increasing_migrants",
         migration_direction = "Increasing")

data_results_increasing_migrants <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsFinalResultsIncreasingMigrationMigrants.txt")) %>% 
  mutate(condition = "Increasing_migrants",
         migration_direction = "Increasing")

```

### Condition: Increasing the frequency of migrations

```{r}
data_timeseries_increasing_frequency <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsIncreasingMigrationFrequency.txt")) %>% 
  mutate(condition = "Increasing_frequency",
         migration_direction = "Increasing")

data_results_increasing_frequency <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsFinalResultsIncreasingMigrationFrequency.txt")) %>% 
  mutate(condition = "Increasing_frequency",
         migration_direction = "Increasing")

```

### Condition: Steady migration, 2 migrants

```{r}
data_timeseries_steady_migration_2_migrants <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsSteadyMigration2Migrants.txt")) %>% 
  mutate(condition = "Steady_migration_2",
         migration_direction = "Steady")

data_results_steady_migration_2_migrants <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsFinalResultsSteadyMigration2Migrants.txt")) %>% 
  mutate(condition = "Steady_migration_2",
         migration_direction = "Steady")


```


### Condition: Steady migration, 5 migrants

```{r}
data_timeseries_steady_migration_5_migrants <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsSteadyMigration5Migrants.txt")) %>% 
  mutate(condition = "Steady_migration_5",
         migration_direction = "Steady")

data_results_steady_migration_5_migrants <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsFinalResultsSteadyMigration5Migrants.txt")) %>% 
  mutate(condition = "Steady_migration_5",
         migration_direction = "Steady")


```


### Condition: No migration

```{r}
data_timeseries_no_migration <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsNoMigration.txt")) %>% 
  mutate(condition = "No_migration",
         migration_direction = "None")

data_results_no_migration <- read.csv(paste0(data_path, "KatsuuraEvaluationResultsFinalResultsNoMigration.txt")) %>% 
  mutate(condition = "No_migration",
         migration_direction = "None")


```

### Testing stuff

```{r}

# data_timeseries_no_migration_test <- read.csv(
#   paste0(data_path, "KatsuuraEvaluationResultsNoMigration.txt")) %>% 
#   mutate(condition = "no_migration",
#          migration_direction = "none")
# 
# data_timeseries_steady_migration_5_test <- read.csv(
#   paste0(data_path, "KatsuuraEvaluationResultsSteadyMigration5Migrants.txt")) %>% 
#   mutate(condition = "steady_migration_5",
#          migration_direction = "steady")
# 
# data_timeseries_combined_test <- data_timeseries_no_migration_test %>% 
#   bind_rows(data_timeseries_steady_migration_5_test)

```



## Combine data

```{r}
data_timeseries_combined <- data_timeseries_increasing_both %>% 
  bind_rows(data_timeseries_increasing_migrants,
            data_timeseries_increasing_frequency,
            data_timeseries_steady_migration_2_migrants, 
            data_timeseries_steady_migration_5_migrants,
            data_timeseries_no_migration) %>% 
  mutate(pretty_condition_newline = str_replace_all(condition, "_", "\n"),
         pretty_condition_space = str_replace_all(condition, "_", " ")) %>% 
  group_by(condition, Run) %>% 
  mutate(is_last_measurement = if_else(Generation == max(Generation), 1, 0)) %>% 
  ungroup()


data_results_combined <- data_results_increasing_both %>% 
  bind_rows(data_results_increasing_migrants,
            data_results_increasing_frequency,
            data_results_steady_migration_2_migrants,
            data_results_steady_migration_5_migrants,
            data_results_no_migration)%>% 
  mutate(pretty_condition_newline = str_replace_all(condition, "_", "\n"),
         pretty_condition_space = str_replace_all(condition, "_", " "))


```


## Analysis

### Timeseries 


```{r}

# Interisland Diversity

tiff("average_interisland_diversity_over_time.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_combined %>% 
  ggplot(aes(Generation, InterIsland.Diversity, color = pretty_condition_space)) +
  geom_smooth() + 
  labs(y="Average interisland diversity",title="Average interisland diversity over time") + 
  theme_bw()
dev.off()

# Global diversity

tiff("global_diversity_over_time.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_combined %>% 
  ggplot(aes(Generation, Global.Diversity, color = pretty_condition_space)) +
  geom_smooth() + 
  labs(y="Global diversity",title="Global diversity over time\nAverage standard deviation of each gene\nacross all individuals") + 
  theme_bw()
dev.off()

# Best score so far

tiff("best_fitness_so_far_over_time.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_combined %>% 
  ggplot(aes(Generation, Best.Score.So.Far, color = pretty_condition_space)) +
  geom_smooth() + 
  labs(y = "Best fitness so far", title="Best fitness so far over time") + 
  theme_bw()
dev.off()

# Number of migrants over time

tiff("migrants_over_time.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_combined %>% 
  filter(condition == "Increasing_both") %>% 
  ggplot(aes(Generation, Migrants)) +
  geom_smooth() + 
  labs(title="Number of migrants over time") + 
  theme_bw()
dev.off()

# Epoch sizes over time

tiff("epoch_size_over_time.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_combined %>% 
  filter(condition == "Increasing_both") %>% 
  ggplot(aes(Generation, Epoch.Size)) +
  geom_smooth() + 
  labs(y="Epoch Size",title="Epoch size over time") + 
  theme_bw()
dev.off()
```

### Final fitness

```{r}

# Box plot of results

tiff("final_fitness_per_condition.tiff", units="in", width=standard_width*1.2, height=standard_height*1.2, res=standard_res)
data_results_combined %>% 
  ggplot(aes(pretty_condition_newline, Final.Score, color = migration_direction)) +
  geom_boxplot()  + 
  #geom_violin() +
  labs(x="Condition", y="Fitness", color="Migration Rate") +
  theme_bw() 
dev.off()
```

Analysis of final results:

```{r}
data_results_combined %>% 
  group_by(pretty_condition_space) %>% 
  summarise(mean_ = mean(Final.Score),
            median_ = median(Final.Score),
            sd_ = sd(Final.Score)) %>% 
  kable()

lm(Final.Score~condition, data=data_results_combined) %>% 
  summary()

```


### Final average global fitness

```{r}

# Average global fitness over time

tiff("global_average_fitness_over_time.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_combined %>% 
  ggplot(aes(Generation, Avg.Global.Fitness, color=condition)) +
  geom_smooth() + 
  labs(y = "Global average fitness", title="The average fitness of all individuals over time") + 
  theme_bw()
dev.off()

# Linear model of avg global fitness by condition
lm(Avg.Global.Fitness~condition, 
   data = data_timeseries_combined %>% 
     filter(is_last_measurement == 1)) %>% 
  summary()

```

### Ranked Islands

Best island? Would maybe say something about initialization?
```{r}
data_timeseries_best_island_combined <- data_timeseries_combined %>% 
  select(c(Run, Generation, condition, Island0, Island1, Island2, Island3, Island4, is_last_measurement, migration_direction, pretty_condition_space, pretty_condition_newline)) %>% 
  gather(Island, AvgIslandFitness, 4:8) %>% 
  mutate(Island = factor(str_replace_all(Island, "Island", ""))) %>% 
  group_by(condition, Run, Island) %>% 
  mutate(is_max_avg_fitness = if_else(AvgIslandFitness == max(AvgIslandFitness), 1,0)) %>% 
  ungroup()
  
# data_timeseries_best_island_combined %>% 
#   ggplot(aes(Generation, AvgIslandFitness, color=Island, fill=condition)) +
#   geom_smooth() + 
#   labs(y = "Average fitness", title="Average Fitness per Island with No Migration") + 
#   theme_bw()

tiff("average_fitness_per_island.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_best_island_combined %>% 
  filter(is_last_measurement == 1) %>% 
  ggplot(aes(Island, AvgIslandFitness, color=pretty_condition_space)) +
  geom_boxplot() + 
  labs(y = "Average fitness", title="Average Fitness per Island") + 
  theme_bw()
dev.off()

tiff("final_average_fitness_per_ranked_island.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_best_island_combined %>% 
  filter(is_last_measurement == 1) %>% 
  arrange(condition, Run, desc(AvgIslandFitness)) %>% 
  group_by(condition, Run) %>% 
  mutate(Rank = factor(1:n())) %>% 
  ggplot(aes(Rank, AvgIslandFitness, color=migration_direction)) +
  geom_boxplot() + 
  labs(x = "Ranked Islands", y = "Final Average Fitness", 
       title="Final Average Fitness per Ranked Island",
       color="Migration Rate") + 
  theme_bw()
dev.off()

tiff("max_average_fitness_per_ranked_island.tiff", units="in", width=standard_width, height=standard_height, res=standard_res)
data_timeseries_best_island_combined %>% 
  filter(is_max_avg_fitness == 1) %>% 
  arrange(condition, Run, desc(AvgIslandFitness)) %>% 
  group_by(condition, Run) %>% 
  mutate(Rank = factor(1:n())) %>% 
  ggplot(aes(Rank, AvgIslandFitness, color=migration_direction)) +
  geom_boxplot() + 
  labs(x = "Ranked Islands",y = "Max. Average fitness", 
       title="Max. Average Fitness per Ranked Island",
       color="Migration Rate") + 
  theme_bw()
dev.off()


```






